---
title: "Proposal"
author: "Robert Lauto"
date: "11/2/2018"
output: html_document
---

### Data Preparation

```{r setup, echo=TRUE, results='hide', warning=FALSE, message=FALSE}
library(jsonlite)
library(httr)
library(tidyverse)
library(curl)
library(dplyr)
library(geniusr)
library(jsonlite)
library(tidytext)
library(assertr)
library(psych)
library(wordcloud)

top_2017 <- read.csv("featuresdf.csv")
spot_daily <- read.csv("spotify's_world_wide_daily_song_ranking.csv")

# Filtering the Spotify data frame down to the US only
usspot = spot_daily %>% 
  filter(Region == 'us' & grepl('2017', Date))

# Calculate the mean rank for each song in the usspot dataframe
mean_rank <- usspot %>% 
  group_by(URL) %>%
  summarise(mean_ranking = mean(Position.x))

# left join to add the mean ranks
merged_w_mranks <- merged_df %>% 
  left_join(mean_rank, by = "URL")

# creating for loop to get top 100 song lyrics

for (i in 1:nrow(top_2017)) {
  
  print('On to the next song')
  print('searching for song in genius')
  sr <- search_song(search_term = as.character(top_2017$name[i]), n_results = 25) %>% 
    filter(grepl(as.character(top_2017$artists[i]), artist_name))
  
  if (is.null(sr$song_id[1]) != T) {
    tryCatch({
      lyricsdf <- scrape_lyrics_id(song_id = sr$song_id[1]) 
     
      print('creating wordcount data frame')
      wordcount <- lyricsdf %>% 
      unnest_tokens(word, line, to_lower = T) %>% 
      count(word, sort = T )
      
      print('creating variables and saving to usspot dataframe')
      max_count <- max(wordcount$n)
      unique_words <- nrow(wordcount)
      max_total_ratio <- max(wordcount$n)/nrow(wordcount)
      lyrics <- tolower(paste(lyricsdf$line, collapse = " "))
      top_2017$lyrics[i] <- lyrics
      top_2017$max_count[i] <- max_count
      top_2017$unique_words[i] <- unique_words
      top_2017$max_total_ratio[i] <- max_total_ratio
    }, error=function(e){cat("ERROR :",conditionMessage(e), "\n")})
  }

  
  Sys.sleep(2)
}
```


### Research question 

**You should phrase your research question in a way that matches up with the scope of inference your dataset allows for.**
  
Is a songs danceability predictive of its popularity?


### Cases 

**What are the cases, and how many are there?**
  
In the daily Spotify dataset, each case is a song that ranked the top 200 most streamed songs for each day in 2017 in the US.There are 72400 observations.
  
In the top_2017 dataset, each case is a song that ranked in Spotify's top 100 most streamed songs for 2017 and each song's audio features. There are 100 observations.


### Data collection 

**Describe the method of data collection.**
  
Data was collected from a playlist Spoify has put together for the 200 most streamed tracks by country each day. Spotify also created a playlist of top 100 songs of 2017 and included descriptive audio features.

### Type of study 

**What type of study is this (observational/experiment)?**
  
This is an observational study.


### Data Source 

**If you collected the data, state self-collected. If not, provide a citation/link.**
  
https://www.kaggle.com/edumucelli/spotifys-worldwide-daily-song-ranking

https://www.kaggle.com/nadintamer/top-tracks-of-2017

Genius.com API

### Dependent Variable

**What is the response variable? Is it quantitative or qualitative?**

The response variable is the ranking or average ranking of a song and it is quantitative.

### Independent Variable

**You should have two independent variables, one quantitative and one qualitative.**
  
The quantitative independent variable is the danceability The qualitative variable is the lyrics that I collect from genius API.

### Relevant summary statistics 

**Provide summary statistics for each the variables. Also include appropriate visualizations related to your research question (e.g. scatter plot, boxplots, etc). This step requires the use of R, hence a code chunk is provided below. Insert more code chunks as needed.**

```{r}
describe(mean_rank$mean_ranking)

describe(top_2017$danceability)

wordcloud(top_2017$lyrics, max.words =100, min.freq = 4, scale = c(4,.5))

plot(density(mean_rank$mean_ranking))

plot(density(top_2017$danceability))
```
